RNA-seq from 4 cell populations

### 2021-11-01 deconvolution/fans4/molec_profiling.Rmd
per cell type

@@

@@

@@

### Esimate precision weights
```{r voom}
form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender) + (1|Batch)
vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
```

>>>>

```{r excludeIDs}
excludeIDs = c('CMC_MSSM_002')
info = info[! info$Individual.ID %in% excludeIDs,]
```

```{r exit2, cache=FALSE}
knitr::knit_exit()
```

@@

# Per cell type analysus
```{r preprocess}

for( CT in unique(info$CellType) ){

  idx = info$CellType == CT
  isexpr = filterByExpr(geneCounts[,idx])

  dge = DGEList( geneCounts[isexpr,] )
  dge = calcNormFactors(dge)

  form = ~ (1|Gender) + (1|Batch)
  vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))

  form = ~ (1|Gender) + (1|Batch) ~ Batch + Gender + Age.of.Death
  fit = dream(vobj, form, info, BPPARAM=SnowParam(12, progressbar=TRUE)))

# contrasts
# L = makeContrastsDream(form, info, contrasts=c(
#               GABA.Dx = "IGABA.SCZ - IGABA.Control",
#               GLU.Dx = "IGLU.SCZ - IGLU.Control",
#               MgAs.Dx = "IMgAs.SCZ - IMgAs.Control",
#               Olig.Dx = "IOlig.SCZ - IOlig.Control"))

}

# Jooint analysis

@@

```{r cancor, eval=FALSE}

pairwise_cca = function(fit){

  exprResid = t(residuals(fit))

  df = apply(combn(4, 2), 2, function(x){

    CT1 = levels(info$CellType)[x[1]]
    CT2 = levels(info$CellType)[x[2]]

    X = exprResid[info$CellType == CT1,]
    Y = exprResid[info$CellType == CT2,]

    idx = match(rownames(X), rownames(info) )
    rownames(X) = info$Individual.ID[idx]

    idx = match(rownames(Y), rownames(info) )
    rownames(Y) = info$Individual.ID[idx]

    tab = table(c(rownames(X), rownames(Y)))
    ids = names(tab[tab==2])

    X = X[ids,]
    Y = Y[ids,]

    mx = rowMeans(geneExpr[,info$CellType == CT1])
    my = rowMeans(geneExpr[,info$CellType == CT2])

    # idx = sample.int(length(ids), length(ids), replace=FALSE)

    X2 = scale(X[,mx>6])
    Y2 = scale(Y[,my>6])

    X2 = apply(X2, 2, RankNorm)
    Y2 = apply(Y2, 2, RankNorm)

    incl = seq(1, 0.8*nrow(X2))

    fit = fastcca(X2[incl,],Y2[incl,])
    X.pred = X2[-incl,] %*% fit$x.coef
    Y.pred = Y2[-incl,] %*% fit$y.coef
    rho.fastcca = cor(X.pred[,1], Y.pred[,1])^2
    rho.fastcca

    data.frame(CT1 = CT1, CT2 = CT2, rho = fit$cramer.V, rho.emp = rho.fastcca)
  })
  df = do.call(rbind, df)

  df
}

dfToMatrix = function(df, V1, V2, value){

  lvls = unique(c(df[,V1], df[,V2]))
  C = diag(1, length(lvls))
  C[C!=1] = NA
  rownames(C) = lvls
  colnames(C) = lvls

  for(i in 1:nrow(df) ){
    C[df[,V1][i], df[,V2][i]] = df[[value]][i]
    C[df[,V2][i], df[,V1][i]] = df[[value]][i]
  }

  C
}

library(decorrelate)

library(RNOmni)

devtools::reload("/hpc/users/hoffmg01/build2/decorrelate")

fit = dream(geneExpr, ~ Batch + Gender + Age.of.Death, info)
fit = eBayes(fit)

C = dfToMatrix(df, "CT1", "CT2", "rho.emp")

plotCorrMatrix(C)

cmd = cmdscale(dist(1-C^2))
df_cmd = data.frame(CellType = rownames(cmd), data.frame(cmd))

ggplot( df_cmd, aes(X1, X2, color=CellType)) + geom_point() + theme_classic() + theme(aspect.ratio=1)

@@

### Estimate precision weights
```{r voom}
form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender) + (1|Batch)
vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(12, progressbar=TRUE))
```
